Quantitative processing of broadband data as implemented in a scientific splitbeam echosounder

### 2022-04-11 Paper/Manuscript.tex
Updated Introduction and Discussion
Direct copy from Word doc on Teams. No edits to text itself yet.

Active acoustics is an efficient method for remote sensing of marine ecosystems. These methods have been efficiently used to cover a wide range of spatial and temporal scales from the aquatic ecosystem, ranging from tens of km \citep{duane_quantification_2021} to small scale behaviour patterns \citep{handegard_avoidance_2003}. These methods are also a crucial component of fisheries resource management. Several different configurations of acoustic instruments are available, these include multibeam sonars, synthetic aperture sonar, omnidirectional sonars, acoustic imaging sonars as well as single- and multibeam echosounders (\hl{refs}).

Single-beam echosounders have extensively been used in fisheries management and ecosystem monitoring. Methodology for using these systems are well developed and in particular the ability for accurate calibration of the system has made these specially useful. Single-beam echosounders have provided important information for fisheries resource management \citep{Simmonds2005Fisheries} as well as insight in ecological processes for the aquatic ecosystem (Godø, 2014).

Conventional echosounders generate an acoustic pulse with a narrow bandwidth (several kHz at most), and the simultaneous operation at widely spaced frequencies (such as 18, 38, 70, 120, 200, and 333 kHz). The method has been used to  study predator-prey interactions \citep{benoit-bird_prey_2003, friedlaender_diel_2009, veit_aggregation_1993}, interactions between marine organisms and the physical environment \citep{crawford_water_2012, freeman_diel_2004, georgiadis_coralligene_2009, ludsin_hypoxia_avoidance_2009} (+Godø, 2014), abundance and distribution of prey and predators \citep{thompson_winter_1991} and non-fish species or group of species \citep{brierley_acoustic_2001}, and the individual or group behavior of marine organisms \citep{axelsen_await_2000, benoit-bird_diel_2001, kaartvedt_use_2009}.

Recent echosounders produce pulses with a wide and continuous frequency range (broadband pulses), as opposed to conventional single frequency systems, i.e. a continuous frequency coverage over a wide frequency band. This provides significantly better along-beam resolution and a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}. These benefits can lead to improved backscatter categorization \citep{korneliussen2018, Traykovski1998Effect, Stanton2012Resonance, Korneliussen2016Acoustic} and hence more accurate distribution and abundance estimates, as well as enhanced information on ecosystem structure and functioning. Early work with recent commercial systems indicate the potential improvement in detection near boundaries \citep{lavery2017} and remote sizing of fish \citep{kubilius_remote_2020}, capabilities of fine-scale observations of interactions between prey and predators \citep{skaret_diel_2020}, the potential for species discrimination \citep{bassett_broadband_2018, blanluet_characterization_2019, benoit-bird_exploring_2020, gugele_differentiation_2021, hasegawa_situ_2021}, and the potential for enhanced information on mesopelagic layers \citep{cotter_classification_2021, agersted_application_2021, khodabandeloo_estimating_2021, khodabandeloo_mesopelagic_2021}.

@@

To tap the full potential of broadband echosounder data, further methods for processing such data must be developed \citep{lavery2017,bassett_broadband_2018}. Examples include adapting pulses to improve target separation as well as further parameters from the individual targets. Different send pulses can be envisioned as well as various methods for acoustic target classification. The objective of this paper is to present a systematic and comprehensive explanation of how to process calibrated broadband echosounder data. By making associated code available, we provide a starting point for further developing signal processing methods for applying broadband acoustic to aquatic ecology.

@@

\section{Discussion}
The use of broadband signals in fisheries acoustics is a developing area and we anticipate many valuable enhancements will occur in the coming years. For example, the use of $\ts(\freqsym)$ and $\sv(\freqsym)$ to improve acoustic target classification \citep{bassett_broadband_2018, korneliussen2018}, and the potential of the high range resolution from pulse compression to observe small-scale fish behviours \citep{skaret_diel_2020} and to detect objects adjacent to boundaries \citep{lavery2017}. The basic formulation for calculating $\ts(\freqsym)$ and $\sv(\freqsym)$ presented here provides the foundation for future enhancements.

>>>>

Active acoustics is an efficient method for remote sensing of marine ecosystems. These methods have been used to cover a wide range of spatial and temporal scales, ranging from tens of km \citep{makris_fish_2006} to small scale behaviour patterns \citep{klevjer_split-beam_2003}, and can provide information on key forage species, including fish and zooplankton, linking primary producers and top predators as well as other taxonomic groups important for the ecosystem functioning \citep{benoit-bird_ecological_2016}. As early as 1935, Oscars Sund observed the distribution of spawning cod in the Lofoten area using a single beam echosounder \citep{sund_echo_1935}. The method was further developed to map the abundance of fish, driven by the need for fisheries management \citep{Simmonds2005Fisheries}. More recently, the methods have been deployed on a wide range of platforms, including observatories, autonomous underwater vehicles \citep{fernandes_autonomous_2003}, uncrewed surface vehicles \citep{de_robertis_uncrewed_2021} and vessels of opportunity, observing a wide range of ecosystem processes across different spatial and temporal scales \citep{godo_marine_2014}.

Several different configurations of acoustic instruments are available, including multibeam sonars \citep{gerlotto_two_1999}, synthetic aperture sonar, omnidirectional sonars \citep{misund_improved_1996}, acoustic imaging sonars \citep{jaffe_ftv_1995} as well as single- and multibeam echosounders \citep{trenkel_new_2008}. Narrowband single beam echosounders have been used extensively in fisheries management and ecosystem monitoring, and methodology for using these systems are well developed. The ability for accurate calibration of the system has made these especially useful for providing data for fisheries management \citep{Simmonds2005Fisheries}.

Recent commercially available single beam echosounders produce pulses with a wide and continuous frequency range (broadband pulses), as opposed to conventional single frequency systems. This provides significantly better along-beam resolution, a higher signal to noise ratio than narrowband pulses \citep{Chu1998Application, ehrenbergFMSlideChirp2000}, and  improved frequency resolution for backscatter categorization \citep{korneliussen2018}.

Broad banded echosounders have been pursued actively for ecology in particularly based on the promise of improved categorization of species or group of species. Using high frequency broadband acoustics (150-600 kHz) Lavery et al \citep{lavery_measurements_2010} showed that broadband helped reduce the ambiguities in interpretation of acoustic scattering from zooplankton and oceanic microstructure. \citep{Stanton2012Resonance} used low frequency broadband pulses (1-6 kHz) with frequency content covering the resonance of swimbladdered fish, demonstrating both the spectral resolution and ability to classify size classes of fish within fish assemblages. \citep{blanluet_characterization_2019} used broadband, with ground truthing, to identify that the composition of two Sound Scattering Layers (SSL) detected in the Bay of Biscay in springtime was characterized and identifying fine scale heterogeneity within the SSLs. While unsuccessful in discriminating between several species of fishes with large  swimbladders during the Alaska pollock survey, \citep{bassett_broadband_2018} showed that broadband may be helpful in characterizing smaller fishes with swimbladder and euphausiids. \citep{benoit-bird_exploring_2020} was able to effectively discriminate three monospecific aggregations of species (hake, anchovy and krill) using broadband data (45-170 kHz), while attempts to classify using narrowband multifrequency failed.

The increased spatial resolution is another aspect of broadband acoustics that are actively being used both for improved categorization and increased ability to resolve targets near each other or near boundaries such as the seafloor. \citep{lavery2017} explored different broadband pulse shapes to increase successful resolution of single targets in proximity as well as near boundaries in tank experiments. By applying high frequency broadband pulses to fish-like artificial targets, \citep{kubilius_remote_2020} demonstrated the potential for acoustic sizing of individually resolved fishes in a controlled ex situ environment. Utilizing the increased spatial resolution \citep{hasegawa_situ_2021} were able to isolate single fishes and discriminate successfully between average frequency responses of walleye pollock and pointhead flounder.

An equally interesting outcome from the increased resolution and broad frequency bands is the ability to infer behavioural information on both single individuals, groups of individuals, and interaction between individuals or groups. \citep{Traykovski1998Effect} was able to extract orientation information about krill from using a combination of high frequency broadband acoustics and acoustic modelling. Fine-scale observations of single targets facilitated the study of interactions between prey and predators, capelin and cod, in the Barents sea (\citep{skaret_diel_2020}).

@@

The theoretical foundation for signal processing comes from radar applications and is earlier documented for broadband echosounders \citep{stanton2008}. To tap the full potential of broadband echosounder data, further data processing methods are being developed \citep{lavery2017,bassett_broadband_2018}. Examples include adapting pulses to improve target separation as well as further parameters from the individual targets. Different send pulses can be envisioned as well as various methods for acoustic target classification. When translating equations into code, several choices must be made. Some of these are well founded in signal processing literature whereas others are of a more practical and ad-hoc nature. The latter is typically missing in conventional papers, and this makes it difficult to test the implementation of the signal processing methods in new echo sounders and post processing software.

The objective of this paper and associated code is to present a systematic and comprehensive description of the data processing steps for calibrated broadband echosounder data. The steps include pulse compression, target strength as a function of frequency [TS(f)], and volume backscattering strength as a function of frequency [Sv(f)]. The intention is that the code will be used to validate implementations in various echosounders and post processing software, provide a starting point for further developing signal processing, and serve as a learning resource for broadbanded acoustic data processing. We also note that the processing equations and methodology presented in this paper are similar to version 1.12.4 and earlier of the Simrad EK80 software.

@@

\section{Discussion}

The use of broadband signals in fisheries acoustics is a developing field, and our contribution represents a comprehensive description of data processing steps. The inclusion of a corresponding code ensures a complete description. One could also envision that signal processing steps designed for specific purposes beyond these standard parameters can be developed, and our code may then serve as a starting point for further development. We also envision that the paper and code will be used both for educational purposes and for understanding the standard signal processing.

Choices

The contribution includes all the minor details required for the processing steps to work in practice. This includes steps well founded in the literature as well as practical and more ad-hoc choices, and a few examples of such choices follows. How to handle gaps in the calibration data, the choice of transmit pulse including tapering, calculation of efficient pulse duration, and decimation factors and filtering are examples of such choices. When convolving the received signal with the transmit pulse, there are various approaches to handle edge cases, and, in our implementation, we simply chose to remove the edge cases. We have also assumed a four-sector transducer, and the code must be adapted to other beam configurations if other configurations are needed. The signal processing also assumes a matched load impedance for the transducer and transceiver. Our objective is not to provide an evaluation of these choices, but to serve as a reference implementation when benchmarking the choices.

$\ts(\freqsym)$ is a common metric for studying single targets, used to extract features for single individuals. These features includes size, target classification and is used to extract behaviour through tracking. In our implementation, the target strength by frequency calculation assumes that a single target has successfully been identified. For this to work in proactive a robust single target detector is required. Several single echo detection algorithms exist, and various single echo detection algorithms may be required depending on situation. Typical SED algorithms are based on traditional single frequency pulses, and utilizing the information in the broadband data improved SEDs may be envisioned, but this is outside the scope of this paper.

$\sv(\freqsym)$ is a key parameter for echo integration. To estimate $\sv(\freqsym)$ a Fourier transform is used, repeatedly applied via a sliding window in range. The recommended size of the window is two times the pulse length, and is chosen as a compromise between spatial and frequency resolution. Since the duration of the sliding window can cause the spreading loss compensation to differ between the beginning and end of the window, the compensation for spreading loss is performed on the pulse compressed time domain data before the transform. Absorption loss compensation is also range dependent (and frequency dependent), but is insignificant for typical operating frequencies at short ranges and between the beginning and the end of the window at longer ranges. The compensation for absorption loss is therefore performed after applying the discrete Fourier transform. The choice of window also allow the data to be cast on a predefined range-frequency grid, which can be used to fit data across transducers to an n-dimensional tensor typically employed by machine learning frameworks like Keras, Pytorch, and Tensor flow.

The formulation presented in this paper results in several frequency dependent parameters, such as transducer gain, two-way equivalent beam angle, and the water absorption coefficient, that are required to quantitatively estimate $\ts(\freqsym)$ and $\sv(\freqsym)$ from received broadband signals. Methods to estimate these are not within the scope of this paper, but common practise is to use the conventional sphere backscatter calibration methodology \citep{demerCalibrationAcousticInstruments2015} slightly enhanced for broadband \citep{hobaekCharacterizationTargetSpheres2013,lavery2017}. We note that these methods do not provide an operational method to estimate $\teff$ or $\Phi(f)$, especially for ship-mounted transducers, and that empirical measurements of these parameters are necessary to fully calibrate both narrowband and broadband echosounders.

A set of equations for calculating calibrated, frequency-dependent, target strength and volume backscatter from broadband echosounder signals have been presented. Providing an open-source resource for anyone interested in learning and developing the techniques further.

Old Discussion below

The use of broadband signals in fisheries acoustics is a developing area and we anticipate many valuable enhancements will occur in the coming years. For example, the use of $\ts(\freqsym)$ and $\sv(\freqsym)$ to improve acoustic target classification \citep{bassett_broadband_2018, korneliussen2018}, and the potential of the high range resolution from pulse compression to observe small-scale fish behviours \citep{skaret_diel_2020} and to detect objects adjacent to boundaries \citep{lavery2017}. The basic formulation for calculating $\ts(\freqsym)$ and $\sv(\freqsym)$ presented here provides the foundation for future enhancements.
